---
title: "project3"
---

Who is doing what research question:

Tiannah - During which time period did hospitalizations peak and how have recovery rates changed over time?

Ruixin - Which countries had the highest cumulative deaths?

Ethan - What is the relationship between positivity and testing rates across different regions and how does this relationship change throughout phases of the pandemic

Ashley - How do hospitalization rates compare across countries?

Kevin - Is there a relationship between hospitalization lag with identification of positive COVID cases, and how does this lag correlate with healthcare infrastructure of a given region

? - To what extent does geolocation contribute to the trajectories of COVID cases in a specific region. Do neighboring countries exhibit similar trends and what factors contribute to these patterns.

? - Make slides, 1 for each question

```{r}
library(ggplot2)
library(plotly)
library(dplyr)

covid_data <- read.csv("tested_worldwide.csv") %>% 
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))
```

```{r}
daily_positive_tested <- covid_data %>%  
  group_by(Country_Region) %>% 
  summarize(
    avg_positive = mean(daily_positive, na.rm = TRUE),
    avg_tested = mean(daily_tested, na.rm = TRUE)
  )

positive_tested_chart <- ggplot(
  daily_positive_tested %>%
    filter(
      !is.na(avg_positive) & !is.na(avg_tested),
      avg_positive > 0,
      avg_tested > 0
    ),
  mapping = aes(
    x = avg_positive,
    y = avg_tested,
    text = paste("Country:", Country_Region)
  )) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "Average Daily Positive Cases vs Quantity of Testing",
    x = "Average Daily Positive Cases",
    y = "Average Daily Quantity of Tests"
  ) +
  theme_minimal()

positive_tested_interactive <- ggplotly(positive_tested_chart)
positive_tested_interactive
```
